---
title: "R4SPISE2024"
format:
  html:
    code-fold: true
    code-summary: "R code"
---

Here, you will find data, programs, and documentation for the SPISE2024 Workshop *A multi-dimensional approach using MCA, MFA, and PLS with R* (from July 24 to July 25, Ho Chi Minh City).

## Reproducible examples in R

Students who are proficient in R are invited to check our reproducible examples out: all the data and code is available on this companion website.

All the necessary code to generate the figures and tables "hidden" in collapsed chunks of codes, like this one below.

Click on the arrow to see the R code.

```{r example R code, eval = FALSE}
# This is a comment!
# And this, below, is a line of R code.
# Copy and paste it in your own R console
# to compute the median of three numbers: 
# -1, 0, and 1
median(-1:1)
# (the result will and should not surprise you)
```


### Install our companion package

We made the data, as well as cool functions to plot your results, available in an R package that you can dowload and install with `pak`.

```{r install, eval = FALSE, collapse=FALSE}
#| code-fold: false
## If needed, install pak first
install.packages("pak")
## Install R4SPISE2022
pak::pak("HerveAbdi/R4SPISE2022")
```

The ouput that you will get will depend on what is already installed on your computer, but should look like this:

```
→ Will install 1 package.
→ Will download 1 package with unknown size.
+ R4SPISE2022   0.0.0.9000 👷🏿️🔧 ⬇ (GitHub: --------)
ℹ Getting 1 pkg with unknown size
✔ Got R4SPISE2022 0.0.0.9000 (source) (7.85 MB)             
✔ Downloaded 1 package (7.85 MB) in 21.8s                   
ℹ Packaging R4SPISE2022 0.0.0.9000
✔ Packaged R4SPISE2022 0.0.0.9000 (7.4s)                    
ℹ Building R4SPISE2022 0.0.0.9000 
✔ Built R4SPISE2022 0.0.0.9000 (2.8s) 
```

### Test

To test the installation, run the following chunk of code that will show the help file of function `OTAplot`.

```{r test installation}
library(R4SPISE2022)
?OTAplot
```


